---
title: "Comparing two networks' centrality indices in the same plot"
subtitle: ""
excerpt: "Creating a function to compare centrality indices of two networks"
author: "Gabriel R. R."
date: 2021-10-11
output: 
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: true
    df_print: paged
    code_folding: show
draft: false
images:
series:
tags:
categories:
- tutorials
- function
- centrality
- qgraph
- network psychometrics
layout: single # or single-sidebar
---
## Getting our two networks

A problem we might encounter when comparing two networks is how to report
centrality indices for these two networks. In this quick tutorial, I created
a function to plot centrality indices differing networks by its graph linetype.

First, let's create two networks based on gender using the BFI dataset from the
*psych* package.

```{r Estimating networks, message=FALSE, warning=FALSE}
load_libraries <- function(){
  if (!require("bootnet"))
    install.packages("bootnet"); library(bootnet)
  if (!require("dplyr"))
    install.packages("dplyr"); library(dplyr)
  if (!require("magrittr"))
    install.packages("magrittr"); library(magrittr)
  if (!require("psych"))
    install.packages("psych"); library(psych)
  if (!require("qgraph"))
    install.packages("qgraph"); library(qgraph)
}

load_libraries()

# Data:
df <- bfi[,1:26]

df$gender %<>% factor(levels = 1:2,
                      labels = c("Male", "Female"))

# Estimating networks:
network_male <- estimateNetwork(df %>% 
                                  filter(gender == "Male") %>% 
                                  select(-gender),
                                default = "EBICglasso",
                                corMethod = "spearman")

network_female <- estimateNetwork(df %>%
                                    filter(gender == "Female") %>% 
                                    select(-gender),
                                  default = "EBICglasso",
                                  corMethod = "spearman")

```


## Comparing two centrality indices

Now that we have our two networks, we could try and compare centrality indices.
The function below could help with that:

```{r compareCentrality function, warning=FALSE,message=FALSE}
compareCentrality <- function(net1, net2,
                              include = c("Strength",
                                          "Closeness",
                                          "Betweenness",
                                          "ExpectedInfluence",
                                          "all",
                                          "All"),
                              orderBy = c("Strength",
                                          "Closeness",
                                          "Betweenness",
                                          "ExpectedInfluence"),
                              decreasing = T,
                              legendName = '',
                              net1Name = 'Network 1',
                              net2Name = 'Network 2'){
  
  library(ggplot2)
  library(forcats)
  
  if(include == "All" | include == "all"){
    include = c("Strength",
                "Closeness",
                "Betweenness",
                "ExpectedInfluence")
  }
  
  df <- centralityTable(net1, net2) %>% filter(measure %in% include)
  
  df %>% 
    mutate(graph = case_when(graph == 'graph 1' ~ net1Name,
                             graph == 'graph 2' ~ net2Name),
           graph = as.factor(graph),
           node = as.factor(node)) %>% 
    
    mutate(node = fct_reorder(node, value)) %>% 
    
    ggplot(aes(x = node, y = value, group = graph)) +
    
    geom_line(aes(linetype = graph), size = 1) +
    
    labs(x = '', y = '') +
    
    scale_linetype_discrete(name = legendName) +
    
    coord_flip() +
    
    facet_grid(~measure) +
    
    theme_bw()
  
}
```

Now that we created the function, let's use all its arguments and see what it
does.

```{r using compareCentrality, message=FALSE, warning=FALSE}
compareCentrality(network_male, network_female,
                  include = "all",
                  legendName = "Networks by Gender",
                  net1Name = "Male",
                  net2Name = "Female")
```

There we go! Now we have plotted our centrality measures on both networks at
the same time :)
