---
title: "Sum specific columns by rows"
subtitle: ""
excerpt: "Sum scores using dplyr::select()"
author: "Gabriel R. R."
date: 2021-08-21
output: 
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: true
    df_print: paged
    code_folding: show
draft: false
images:
series:
tags:
categories:
- dplyr
- tutorials
- function
layout: single # or single-sidebar
---
## Sometimes you have a messy dataset

By that, I mean a dataset with a messy column ordering, uneccessary variables
and so on. That's somewhat the case with the DASS-42 dataset taken from
Kaggle (available 
[here](https://www.kaggle.com/lucasgreenwell/depression-anxiety-stress-scales-responses)).

I've taken the liberty of preserving "only" 10k cases. Let's load the data and
present the problem.

```{r Loading data and package, include=T, warning=F, message=F}
load_libraries <- function(){
  if (!require("dplyr"))
    install.packages("dplyr"); library(dplyr) # select() and mutate()
  if (!require("ggplot2"))
    install.packages("ggplot2"); library(ggplot2) # final plot
  if (!require("magrittr"))
    install.packages("magrittr"); library(magrittr) # %<>% operator

}

load_libraries()

df <- read.csv('https://raw.githubusercontent.com/GabrielReisR/R/master/estrutura%20de%20dados/dados/dass42_brief.csv')[-1]

names(df)
```

The thing here is: item responses are stores in the "Q_number_A" columns.
And columns are ordered in a weirdy weird way.
The DASS scale measures depression, anxiety and stress. Say we're interested
in joining these three constructs in an umbrella construct called *neuroticism*.

In order to do that, we'd need to sum all items that start with a Q and ended
with an A. We can do that with *dplyr*.

```{r Solving problem}
df %<>% 
  mutate(neuroticism = df %>% select(starts_with("Q") & ends_with("A")) %>%
                  rowSums())
```

What we've done is:

1. Created a column called *neuroticism* using the `mutate` function.
2. Selected all variables in *df* that started with "Q" and ended with an "A".
3. Used `rowSums()` on those variables and stored it in *neuroticism*.

Let's check out our new variable!

```{r Histogram of neuroticism scores}
df %>%
  ggplot(aes(neuroticism)) +
  geom_histogram(aes(y = ..density..), bins = 40,
                 colour = "#011e5a", fill = "white") +
  stat_function(fun = dnorm,
                args = list(
                  mean = mean(df$neuroticism, na.rm = T),
                  sd = sd(df$neuroticism, na.rm = T)),
                colour = "#011e5a") +
  
  # Rótulos
  xlab('Neuroticism Score') +
  ylab('Density of Scores') +
  
  # Tema
  theme_classic()
```

It worked!
That's it for today!
