---
title: "Estimating a psychometric network with qgraph"
subtitle: "Learn how to estimate a psychometric network with qgraph"
excerpt: "Understand qgraph's simplicity in estimating psychometric networks."
author: "Gabriel R. R."
date: 2021-08-10
output: 
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: true
    df_print: paged
    code_folding: show
draft: false
images:
series:
tags:
categories:
- network psychometrics
- tutorials
- qgraph
layout: single # or single-sidebar
---
## We're gonna need data

To estimate a psychological network we're going to need data.
Let's take the 25 item Big-Five questionnaire from the *psych* package.

```{r Initial setup, include=T, warning=F, message=F}
load_libraries <- function(){
  if (!require("dplyr"))
    install.packages("dplyr"); library(dplyr)
  if (!require("psych"))
    install.packages("psych"); library(psych)
  if(!require("qgraph"))
    install.packages("qgraph"); library(qgraph)
}

load_libraries()

df <- bfi[,1:25]
glimpse(df)
```

Note that we have ordinal data. This is not problematic and we'll mostyl deal
with it using the `cor_auto` function from *qgraph*. The `cor_auto` function
estimates a correlation matrix with the appropriate method for our data. This
is done by automatically (hence the name) identifying our data type.

We have now three initial alternatives of estimation using the *qgraph* package:

1. Estimate network with a correlation matrix
2. Estimate network with a partial correlation matrix
3. Estimate network using EBICglasso

Before going into the actual code, let's prepare this hyperparameter that'll
separe our nodes by groups.

```{r Traits}
traits <- rep(c('Agreeableness',
                'Conscientiousness',
                'Extraversion',
                'Neuroticism',
                'Openness'),
              each = 5)
```

## 1. Estimate network with a correlation matrix

```{r Estimate network with a correlation matrix, warning=F, message=F}
network <- qgraph(cor_auto(df),
                  graph = 'cor',
                  layout = 'spring',
                  theme = 'colorblind',
                  groups = traits)
```

The correlation network plots the bivariate association between two variables.
In this sense, when estimating an edge between two variables, it does not take 
into account the covariance existing between other variables. In my 
understanding, a graph like this would probably be used to show the bivariate 
association between a small number of variables.

## 2. Estimate network with a partial correlation matrix

```{r Estimate network with a partial correlation matrix, message=F}
network <- qgraph(cor_auto(df),
                  layout = 'spring',
                  graph = 'pcor', # note this new argument
                  theme = 'colorblind',
                  groups = traits)
```

One way to take into account other variables' influence while estimating the
unique association between two variables is by estimating a partial correlation
matrix. This can be done with the `graph = 'pcor'` argument.

A graph that plots the unique association between two nodes is called a 
**Gaussian Graphical Model** (GGM). According to Burger et al. (in press), in a
GGM the parameters are representing the unique association among two variables
after conditioning on all other variables in the network.

## 3. Estimate network using EBICglasso
Note that the graph above is plotting the unique association between every
two variables. A possible problem with the graph above is that we're possibly
plotting false positives since any partial correlation different than 0 is being
shown. In that sense, a correlation of 0.02 would be plotted following the
procedure above. The thing is: spurious correlations may not be present in the 
true model - they're probably Type I errors.

We can prune and threshold our network to fix that, but for now I'd like to
present another form of estimating a GGM while controling for Type I errors.
This method is called EBICglasso - Extended Bayesian Information Criteria after
estimating a graphical LASSO network (Foygel & Drton, 2010).

The code is fairly simple and, as you can see, the network is sparser (i.e.,
the network has less parameters, or edges). The only new argument should be
the sample size used in estimating the variance-covariance matrix. This could
be a number `sampleSize = 2800` or you can directly pull this
number from the dataframe (if you don't have missing values),
`sampleSize = nrow(df)`.

```{r Estimate network using EBICglasso, message=F, warning=F}
network <- qgraph(cor_auto(df),
                  layout = 'spring',
                  graph = 'glasso',
                  sampleSize = nrow(df), # new argument!
                  theme = 'colorblind',
                  groups = traits)
```
Note that it's sparser! How does it achieve that? The EBICglasso follows these
steps:

1. Estimate partial correlation network *S*
2. Regularize *S* with the LASSO penalization *n* times using *n* different 
numbers of λ varying from 0 to 1.
3. Among the *n* networks, use the EBIC criterion to select the final network.

One important thing is that the EBIC uses a parameter called γ, that varies from
0 to 1. A value of γ = 0 aims toward discovery and allows for more Type I 
errors. A value of γ = 1 is conservative and will probably commit more Type II
errors. The recommended value of γ is 0.5 according to Epskamp and Fried (2018).
This is also the default in *qgraph*.

## References
Burger, J., Isvoranu, A. M., Lunansky, G., Haslbeck, J. M. B., Epskamp, S., 
Hoekstra, R. H. A., Fried, E. I., Borsboom, D., & Blanken, T. F. (in press).
Reporting standards for psychological network analyses in cross-sectional data.
https://psyarxiv.com/4y9nz/

Costantini, G., Epskamp, S., Borsboom, D., Perugini, M., Mõttus, R., Waldorp, 
L. J., & Cramer, A. O. J. (2015). State of the aRt personality research: 
A tutorial on network analysis of personality data in R. 
*Journal of Research in Personality*, *54*, 13–29. 
https://doi.org/10.1016/j.jrp.2014.07.003

Epskamp, S., Cramer, A. O. J, Waldorp, L. J., Schmittmann, V. D., Borsboom, D. 
(2012). qgraph: Network visualizations of relationships in psychometric data. 
*Journal of Statistical Software*, *48*(4), 1–18. 
https://doi.org/10.18637/jss.v048.i04 

Epskamp, S., & Fried, E. I. (2018). A tutorial on regularized partial 
correlation networks. 
*Psychological Methods*, *23*(4), 617–634. 
https://doi.org/10.1037/met0000167

Foygel, R., & Drton, M. (2010). Extended Bayesian Information Criteria for 
gaussian graphical models. 
*Proceedings of the 23rd International Conference on Neural Information Processing Systems*, 604–612. arxiv.org/pdf/1011.6640.pdf
