---
title: "Estimating a psychometric network with qgraph"
subtitle: "Learn how to estimate a simple network with qgraph"
excerpt: "Understand qgraph's simplicity in estimating networks."
author: "Gabriel R. R."
date: 2021-08-07
output: 
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: true
    df_print: paged
    code_folding: show
draft: false
images:
series:
tags:
categories:
- network psychometrics
- tutorials
layout: single # or single-sidebar
---
## A lot can be done with just a matrix

We're gonna learn how to estimate a simple network with qgraph.
For starters, we're gonna need to install and/or load some packages.

```{r Initial setup, include=T, warning=F, message=F}
load_libraries <- function(){
  if (!require("dplyr"))
    install.packages("dplyr"); library(dplyr)
  if(!require("qgraph"))
    install.packages("qgraph"); library(qgraph)
}
load_libraries()
```

There we go! Now, say we have a symmetrical matrix called *mat*. 
This matrix could be interpreted as a mock-correlation matrix. To be fairly
simple, we only need a matrix to estimate a network in qgraph.

```{r Creating matrix}
mat <- matrix(
  c(
       0,  0.3,    0, -0.3, 0.2, 0.3,
     0.3,    0, -0.9,    0,   0,    0,
       0, -0.9,    0,  0.8,   0,    0,
    -0.3,    0,  0.8,    0, 0.3,    0,
     0.2,    0,    0,  0.3,   0,    0,
     0.3,    0,    0,    0,   0,    0
    ),
  ncol = 6, nrow = 6,
  byrow = TRUE)

network <- qgraph(mat)
```

## Accessing layout

Ok. I think one important thing to understand is how qgraph stores the layout
of the network. This is helpful, for instance, if we want to compare networks.

The layout is a rectangular matrix of Xs and Ys for each node. Let's change the 
values of *mat*, get the layout from our initial *network* and then plot
the network again, keeping the same exact structure from before.

```{r Keeping layout}
mat <- matrix(
  c(
       0,  0.9,    0,  0.5, 0.2, -0.1,
     0.9,    0, -0.4,    0,   0,    0,
       0, -0.4,    0,  0.2,   0,    0,
     0.5,    0,  0.2,    0, 0.3,    0,
     0.2,    0,    0,  0.3,   0,    0,
    -0.1,    0,    0,    0,   0,    0
    ),
  ncol = 6, nrow = 6,
  byrow = TRUE)

qgraph(mat, layout = network$layout)
```

We can access the layout object using `network$layout`. We can use this to
change the nodes' location directly. But that isn't common nor very useful
(except if you're plotting a network simulation).

## Saving network

The last initial step comprises saving our network in the correct way. We can't
rely on RStudio's Plot window because the network will try to adapt to that
window's size. Instead, we save our network using the arguments `filetype` and
`filename`. That way, our network gets saved in our current directory. You can
also access a nested directory directly in `filename`.

```{r Saving network, eval=F}
qgraph(mat, 
       layout = network$layout,
       filetype = 'png', # can be 'R', 'pdf', 'svg', 'tex', 'jpg', 'tiff', 'png'
       filename = '\\figures\\network') # opens 'figures' paste and saves there
```

It's important to note that, if we're dealing with a network that has a legend,
we must specify width to be 1.4 times the height.

That's it for now :)
